# Preloaded data and `mtcars`

R comes with several built-in data sets, **which are generally used as demo data for playing with R functions.**

To see the datasets type:

`mtcars`


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Practicing with `mtcars` data set

This demonstration is based on the datasset `mtcars`.

1. Read in `mtcars`
```{r}
data(mtcars)
```


2. View first few rows and last few rows of mtcars dataframe using functions `head()` and `tail()`
```{r}
head(mtcars)
tail(mtcars)
```



3. Some info about mtcars dataframe using function `colnames()`, `rownames()`, `summary()`and `dim()`
```{r}
colnames(mtcars)
rownames(mtcars)
summary(mtcars)
dim(mtcars)
```


4. To calculate the variance of weight:

```{r}
var(mtcars$wt)
```

5. To get the histogram of hp, the code below will produce a histogram:

```{r}
hist(mtcars$hp)
```

6. To calculate the quantiles by percent:

```{r}
quantile(mtcars$wt, c(.2, .4, .8))

```





## Excerises for you:

1. Find the minimum and maximum value of `mpg` 
2. Find the mean and standard deviation of data variable `mpg` 
3. What variable has a 3rd quartile value of 180.0? 

4. Create and explain what this means

```{r, echo=FALSE}
plot(mtcars)
```

5. Create and explain what this means

```{r, echo=FALSE, fig.dim = c(3, 2)}
cor(mtcars)
```

6. Create a variable called `efficiency` which is mpg divided by weight.  Which car has the max `efficiency` and what is this value?
7. Which variable in this dataset has the greatest standard deviation?
8. How many cars have 3 gears?
9. How many cars get more than 17 mpg?
