# Preloaded data and `mtcars`

R comes with several built-in data sets, **which are generally used as demo data for playing with R functions.**

To see the datasets type:

`mtcars`


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Practicing with `mtcars` data set

This demonstration is based on the datasset `mtcars`.

1. Read in `mtcars`
```{r}
data(mtcars)
```


2. View first few rows and last few rows of mtcars dataframe using functions `head()` and `tail()`
```{r}
head(mtcars)
tail(mtcars)
```



3. Some info about mtcars dataframe using function `colnames()`, `rownames()`, `summary()`and `dim()`
```{r}
colnames(mtcars)
rownames(mtcars)
summary(mtcars)
dim(mtcars)
```


4. To calculate the variance of weight:

```{r}
var(mtcars$wt)
```

5. To get the histogram of hp, the code below will produce a histogram:

```{r}
hist(mtcars$hp)
```

6. To calculate the quantiles by percent:

```{r}
quantile(mtcars$wt, c(.2, .4, .8))

```

## Excerises for you:

1. Find the minimum and maximum value of `mpg` 
2. Find the mean and standard deviation of data variable `mpg` 
3. What variable has a 3rd quartile value of 180.0? 

4. Create and explain what this means

```{r, echo=FALSE}
plot(mtcars)
```

5. Create and explain what this means

```{r, echo=FALSE, fig.dim = c(3, 2)}
cor(mtcars)
```

6. Create a variable called `efficiency` which is mpg divided by weight.  Which car has the max `efficiency` and what is this value?
7. Which variable in this dataset has the greatest standard deviation?
8. How many cars have 3 gears?
9. How many cars get more than 17 mpg?

# More simple data wrangling

## a matirx

```{r, echo=FALSE}
x <- matrix(1:15,nr=3)
x
```



1. Write the code that creates this matrix:

2. Write DIFFERENT code that creates this matrix in an alternate way:

3. In the matrix above, what does `[,4]` mean?

4. What code would return the value in the 3rd column and 3rd row?

5. What single line of would give you the average of the all the numbers in columns 2, 4, and 5 and in rows 1 and 3?

6. turn `x` into a data frame.  

7. How do you **know** you have turned `x` into a data frame?

## More fun

```{r, echo=FALSE}
df <- data.frame(matrix(1:100, nr=10))

```


```{r}
df
```

1.  Consider the dataframe above called `df`.  What would running this code return `sum(df[7,7:10])`

```{r, echo=FALSE, eval=FALSE}
df <- data.frame(matrix(1:100, nr=10))
sum(df[7,7:10])
```

2. How can you tell if an object in R is a dataframe?


3. How could you create the dataframe above called `df`?


4. What code would return the average of row 2 of `df`?


5. Consider `mtcars` dataset that comes preloaded with R that looks like this:
```{r out.width="50%"}
head(mtcars)
```

6. Why do I get this error when I run the code below: `Error in plot(hp, mpg) : object 'hp' not found`?

```{r eval=FALSE}
plot(hp,mpg)
```

`Error in plot(hp, mpg) : object 'hp' not found`

Bonus: What is a topic that you find confusing at this point in class? 

## open the `women` data set 
 
1. Add a another variable (column) to the `women` dataframe called GPA which is these 15 numbers: 1.5, 3.7, 4,1, 3, 2.5, 3.8, 0.8, 2, 4, 1, 3, 2.5, 3.0, 4.0.  You shoud get something that looks similar to mine.


```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=FALSE}
data(women)
women$GPA <- c(1.5, 3.7, 4,1, 3, 2.5, 3.8, 0.8, 2, 4, 1, 3, 2.5, 3.0, 4.0)
library(dplyr)
dplyr::tibble(women)
```

2. Use GPA and weight to predict the height of a person who is 155 pounds and has a GPA if 3.33. What is your prediction? 

3. Is GPA a significant predictor of height and how do you know? 

4. Create a figure showing a best fit line on of height and GPA.  

5. Install the dplyr package into your Rstudio session.
